Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2022-12-21 paper-1/manuscript.tex
Rev Fig1

When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons). Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}. In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
\begin{align}
    v_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
    \tau_{\theta} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
\end{align}
with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Pozzorini2015-ei}.

\begin{figure*}%
    \centering
    \includegraphics{fig1.pdf}
    \caption{}
    \label{fig1}
\end{figure*}
\textbf{Fig. 1}Model and paradigm. \textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron, illustrated above the plots, firing pre-determined spikes. The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing. Top: Membrane voltage (black), firing threshold baseline (gray) and adaptive threshold (green) demonstrating TA. Middle: Excitatory ($g_e$, blue) and inhibitory ($g_i$, red) synaptic conductances, demonstrating excitatory STD. Bottom: Inputs to $g_e$ and $g_i$ (vertical bars in the appropriate colors), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line). \textbf{b} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from two neurons of each type. \textbf{c} Distance from stimulation site 0 in terms of minimum number of excitatory synapses. Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (round markers), which are regularly spaced on a circle with a radius of 2.5 mm.

>>>>

When a neuron's membrane potential reached the firing threshold $V_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons). Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $V_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}. In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
\begin{align}
    V_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
    \tau_{\theta} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
\end{align}
with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Pozzorini2015-ei}.

\begin{figure*}%
    \centering
    \includegraphics{fig1.pdf}
    \caption{}
    \label{fig1}
\end{figure*}
\textbf{Fig. 1} Model and paradigm. \textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes, illustrated above the plots. The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing. Top: Inputs to the (post-)synaptic conductances $g_e$ and $g_i$ (vertical bars), and the STD depression variable $x_{Exc}$ of the excitatory presynaptic neuron (dashed line). Note that excitatory inputs are equal to $U x_{Exc} w$ as per Equation \ref{eqn-gsyn}. Middle: Excitatory and inhibitory synaptic conductances ($g_e$ and $g_i$, respectively). Bottom: Membrane voltage, firing threshold baseline (gray) and adaptive threshold, demonstrating TA in the excitatory postsynaptic neuron. \textbf{b} Sample network layout, showing the spatial location of excitatory and inhibitory neurons, and all outgoing synaptic connections from two neurons of each type. \textbf{c} Distance from stimulation site 1 in terms of minimum number of excitatory synapses. Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (large circles), which are regularly spaced on a circle with a radius of 2.5 mm. \textbf{d} Illustration of the stimulation paradigm. Each marker represents a stimulation, with the stimulus identity indicated by color. The target stimulus (A) is represented as solid red, and non-target stimuli are represented with faded colors. Three randomized sequences were presented, containing 80\% A and 20\% B (std), 20\% A and 80\% B (dev), and 20\% of each of 5 stimuli, including A and B (msc). Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period.
